Pulsar Acceleration Shifts from nearby Supernova Explosion

### 2015-12-23 draft3_TC.tex
added positive tones to the end of discussion
1. I remove the line-of-sight discussion since when \theta is small, the second acceleration change is small, so there is effectively only 1 acceleration change.
2. I added the possibility to use many pulsar from SKA to jointly search for an acceleration change as soon as an SNe is detected. Since the timing accuracy is 20ps for best timed pulsars, it is not unreasonable to expect something within a decade or two after SNe.
3. I added a sentence of joint timing to the abstract. We might need to discuss what to highlight in the abstract.
4. My compi

The observed value of $(\dot{w_e}/w_e)$ cannot be exactly what we calculated, because in practice, there are many other contributions to the acceleration. For example, they cna be both in the galaxy center which has a relative acceleration to our spiral arm. However, if the pulsar is close enough to the SN, then we can be pretty certain that in the time scale comparable to $\Delta t$, the SN is the dominant cause of acceleration change. % Afterall, there are rarely any event more violent than a nearby SN explosion. And if there is, we probably would have noticed.

@@

In practice, measuring $\dot{w_e}$ is actually letting it first build up a frequency change, and then to build up again in the phase change. That means within the duration of the middle stage, the arrival time of the pulse changes by 
\begin{equation}
\delta T_{\rm arrival} \sim \Delta\left(\frac{\dot{w_e}}{w_e}\right)(\Delta t)^2
\sim \Delta M~.
\end{equation}
Interestingly, the $1/r_p^2$ in gravitational force exactly cancels with the accumulation time $(\Delta t)^2$. That means independent of $r_p$, a pulsar can detect this change as long as the timing accuracy is better than the light-crossing time of the effective Schwarzschild radius of the neutrino shell\footnote{In practice, we still want the pulsar to be somewhere close to the SN. Without that condition, we cannot be certain that there are no other comparable changes to acceleration during $\Delta t$.}. We can estimate $\Delta M\sim$ a fraction of solar mass $\sim10^{-6}$ seconds. $\delta T_{\rm arrival}$ for a stable, millisecond pulsar is already better \cite{PulsarTiming}\footnote{Our conservative estimation assumes only two measurements: in the beginning and the end of the duration $\Delta t$. In practice, one of course repeat $N$ measurements during the process, which increases the accuracy by another factor of $\sqrt{N}$.}.

With the forecast of thousands of millisecond pulsars to be found by the Square Kilometer Array \cite{SKA}, a new SN effectively sits within a timing-grid. At the instant the SNe is detected at earth, even if only from direct neutrino detection, we know that all pulsar timing periodic derivatives change, and by different amounts depending on their distances. The closest or the best timed pulsars may detect such effect within a decade or two, and others gradually follow. Such a joint timing can triangulate the location of the SN and prepare us to detect the second acceleration change. Given a few SNe per century\cite{SNrate06}, an actual detection is in the forewseeable future.

>>>>

The observed value of $(\dot{w_e}/w_e)$ cannot be exactly what we calculated, because in practice, there are many other contributions to the acceleration. For example, both the SN and the pulsar may be in the galactic center which has a relative acceleration to our spiral arm. However, if the pulsar is close enough to the SN, then we can be pretty certain that in the time scale comparable to $\Delta t$, the SN is the dominant cause of acceleration change. % Afterall, there are rarely any event more violent than a nearby SN explosion. And if there is, we probably would have noticed.

@@

In practice $\dot{w_e}$ is measured by allowing the signal to build up a frequency change, and then to build up again in the phase change. That means within the duration of the middle stage, the arrival time of the pulse changes by 
\begin{equation}
\delta T_{\rm arrival} \sim \Delta\left(\frac{\dot{w_e}}{w_e}\right)(\Delta t)^2
\sim \Delta M~.
\end{equation}
Interestingly, the $1/r_p^2$ in gravitational force exactly cancels with the accumulation time $(\Delta t)^2$. That means independent of $r_p$, a pulsar can detect this change as long as the timing accuracy is better than the light-crossing time of the effective Schwarzschild radius of the neutrino shell\footnote{In practice, we still want the pulsar to be somewhere close to the SN. Without that condition, we cannot be certain that there are no other comparable changes to acceleration during $\Delta t$.}. We can estimate $\Delta M\sim$ a fraction of solar mass $\sim10^{-6}$ seconds. $\delta T_{\rm arrival}$ for a stable, millisecond pulsar is already better \cite{PulsarTiming}\footnote{Our conservative estimation assumes only two measurements: in the beginning and the end of the duration $\Delta t$. In practice, one has $N$ measurements during the process, which increases the accuracy by another factor of $\sqrt{N}$.}.

With the forecast of thousands of millisecond pulsars to be found by the Square Kilometer Array \cite{SKA}, a new SN effectively sits within a timing-grid. At the instant the SN is detected at earth, even if only from direct neutrino detection, we know that all pulsar timing periodic derivatives change, and by different amounts depending on their distances. The closest or the best timed pulsars may detect such effect within a decade or two, and others gradually follow. Such a joint timing can triangulate the location of the SN and prepare us to detect the second acceleration change. Given a few SN per century\cite{SNrate06} an actual detection is possible in the foreseeable future.
